### 2025-04-30 docs/progress-report/sections/conclusion.tex

\documentclass[../apG48.tex]{subfiles}

\begin{document}

Through benchmarking across Spring WebFlux, Spring MVC, and Quarkus, we evaluated eight different template engines and found that non-blocking implementations—especially those using virtual threads—consistently deliver performance on par with traditional blocking approaches under high concurrency. 
These results highlight virtual threads as a promising alternative to complex asynchronous programming models, offering a simpler development experience without sacrificing scalability or responsiveness.

The next steps include a deeper analysis of the benchmark results in the final report, as well as expanded coverage of the current state of non-blocking PSSR and asynchronous programming techniques. Additionally, the project may be extended to support other frameworks such as Vert.x.

\end{document}