### 2025-04-30 docs/progress-report/apG48.tex
small changes and fixes; add docs folder

\documentclass[a4paper,twoside,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{subfiles}
\usepackage{placeins}

\setlength{\textheight}{24.00cm}
\setlength{\textwidth}{15.50cm}
\setlength{\topmargin}{0.35cm}
\setlength{\headheight}{0cm}
\setlength{\headsep}{0cm}
\setlength{\oddsidemargin}{0.25cm}
\setlength{\evensidemargin}{0.25cm}
\setlength{\parskip}{0.1cm}

\title{
    \includegraphics[width=80mm]{logoISEL.png}\\[12pt]
    {\fontsize{14pt}{16pt}\selectfont Instituto Superior de Engenharia de Lisboa}\\[12pt]
    {\fontsize{16pt}{18pt}\selectfont \textbf{Non-blocking Progressive Server-side Rendering (PSSR) Benchmark}}\\[6pt]
    {\fontsize{11pt}{16pt}\selectfont \textbf{Progress Report}}\\[6pt]
    {\fontsize{11pt}{16pt}\selectfont Bsc in Computer Science and Engineering}\\[10pt]
}

\author{
\begin{tabular}{c}
    {\fontsize{14pt}{16pt}\selectfont Bernardo Pereira, A50493@alunos.isel.pt, n.º 50493}
\end{tabular}}

\date{
\begin{tabular}{ll}
  {\fontsize{14pt}{15pt}\selectfont Supervisor:} & {\fontsize{14pt}{16pt}\selectfont Miguel Carvalho, miguel.gamboa@isel.pt}
\end{tabular}\\
\vspace{5mm}
April, 2025}

\begin{document}

\maketitle

\vspace{-1cm}

\section{Introduction}

\subfile{sections/introduction.tex}

\section{Benchmark Implementation}

\subfile{sections/benchmark.tex}

\section{Benchmark Results}

\subfile{sections/results.tex}

\section{Conclusion}

\subfile{sections/conclusion.tex}

\bibliographystyle{unsrt}
\bibliography{referencias}

\end{document}