### 2025-05-14 docs/article/Sections/conclusion.tex

\section{Conclusion}

Through benchmarking across Spring WebFlux, Spring MVC, and Quarkus, we
evaluated eight different template engines and found that non-blocking
implementations—especially those using virtual threads—consistently deliver
performance on par with traditional blocking approaches under high concurrency.
These results highlight virtual threads as a promising alternative to complex
asynchronous programming models, offering a simpler development experience
without sacrificing scalability or responsiveness.

\pagebreak