### 2025-05-21 docs/article/Sections/introduction.tex

\section{Introduction}

% Falta adicionar referências

Modern web applications rely on different rendering strategies to optimize
performance, user experience, and \textit{scalability}. The two most dominant
approaches are \textit{Server-Side Rendering (SSR)} and \textit{Client-Side
    Rendering (CSR)}.

SSR generates HTML content on the server before sending it to the client,
resulting in faster initial page loads and better Search Engine Optimization
(SEO). However, SSR can increase server load and reduce
\textit{throughput}\footnote{The number of requests the server can handle per
    second (RPS)} since each request requires additional processing before
responding.

In contrast, CSR shifts the rendering workload to the browser. The server
initially sends a minimal HTML document with JavaScript, which dynamically
loads the page content. While CSR reduces the server’s burden, it can lead to
slower initial load times, as users must wait for JavaScript execution before
meaningful content appears.

\textit{Progressive Server-Side Rendering (PSSR)} combines benefits from both SSR and CSR
by streaming HTML content progressively. This technique enhances user-perceived performance by
allowing immediate rendering while data loads asynchronously. In this respect,
PSSR is similar to CSR in that the server initially sends a minimal HTML
document to the client and subsequently streams additional HTML fragments as data becomes available.
However, unlike CSR, PSSR retains all rendering responsibilities on the server side,
thereby reducing the load on the client. As a result, the client does
not need to execute JavaScript or make additional requests to retrieve page content.
To enable non-blocking PSSR, template engines require support for asynchronous
data models and non-blocking I/O operations. While some template engines, such
as HtmlFlow, and Thymeleaf, have support for these features, traditional
template engines that use \textit{external} domain-specific languages (DSLs)
often rely on blocking interfaces, such as \texttt{Iterable}, to define the
data model. This reliance on blocking interfaces leads to the system thread
being blocked until the entire HTML content is rendered, which can
significantly impact performance and scalability.

Efficient thread utilization is crucial for low-thread servers, necessitating
asynchronous, non-blocking request handling. In these architectures,
asynchronous I/O operations (e.g., database queries, and API calls) allow
servers to manage multiple concurrent requests efficiently by avoiding blocking
threads while waiting for I/O operations to complete. In these architectures,
where idle time during I/O operations is reduced, non-blocking PSSR enhances a
server's capacity to handle high request volumes with limited threads and
immediately send partial responses to clients.

With the introduction of \textit{Virtual Threads} in Java 21, it is now possible to execute blocking
I/O operations in a non-blocking manner, allowing for the use of these 
traditional template engines in non-blocking contexts without the need for
complex asynchronous programming models. This advancement enables the 
implementation of PSSR using familiar HTML syntax, simplifying template
development and adoption.

This work will focus on exploring existing approaches to non-blocking PSSR,
which include the use of \textit{reactive programming} and \textit{coroutines}
to implement non-blocking I/O operations. As an alternative to these
approaches, we will investigate the viability of \textit{Virtual Threads} in Java 21 
for implementing non-blocking PSSR\@. Section 2 provides an overview
of the state-of-the-art in non-blocking PSSR and HTML template engines. Section
3 provides a more detailed description of the issues associated with
traditional template engines for non-blocking PSSR, and how this work aims to
address them. Section 4 describes the benchmark methodology and implementation.
Finally, Sections 5 and 6 present the results and conclusions of this work,
respectively.