### 2025-05-27 docs/article/Sections/state-of-the-art.tex

\section{Background and Related Work}

In this section, we first examine the different design choices adopted by web
servers in their internal architectures, and how these choices impact the
behavior of the web template engines. Then, in Subsection~2.2, we present the
main properties that characterize each web template technology approach, along
with the advantages and drawbacks that result from these characteristics.

\input{Sections/state-of-the-art-web-servers.tex}
\input{Sections/state-of-the-art-web-templates.tex}