### 2025-05-27 docs/article/Sections/state-of-the-art.tex

\section{Background and Related Work}

In this section, we first present the main properties that characterize each web
template technology approach, along with the advantages and drawbacks that result
from these characteristics. Then, in Subsection~2.2, we examine the different
design choices adopted by web servers in their internal architectures, and how
these choices impact the behavior of the web template engines.

\input{Sections/state-of-the-art-web-templates.tex}
\input{Sections/state-of-the-art-web-servers.tex}