### 2025-05-28 docs/article/Sections/benchmark.tex
Section 4: Include approaches from Spring MVC and Quarkus

\section{Benchmark Implementation}

The benchmark\footnote{\url{https://github.com/fmcarvalho/comparing-non-blocking-progressive-ssr}}
is designed with a modular architecture, separating the \textit{view} and
\textit{model} layers from the \textit{controller} layer ~\cite{Bucanek2009},
which allows for easy extension and integration of new template engines and
frameworks. It also includes a set of tests to ensure the correctness of
implementations and to validate the \textit{HTML} output. It includes two
different data models, defined as \texttt{Presentation} and \texttt{Stock}, as
shown in \autoref{lst:stock-class} and \autoref{lst:presentation-class},
following the same model as
\textit{spring-webflux-comparing-template-engines}~\footnote{\url{https://github.com/xmlet/spring-webflux-comparing-template-engines}}.
The \texttt{Presentation} class represents a presentation with a title, speaker
name, and summary, while the \texttt{Stock} class represents a stock with a
name, URL, symbol, price, change, and ratio.

\begin{center}
      \begin{minipage}{0.48\textwidth}
            \begin{lstlisting}[
  language=kotlin,
  numbers=none,
  basicstyle=\scriptsize\ttfamily,
  caption={\texttt{Stock} class},
  label={lst:stock-class},
]
data class Stock(
    val name: String,
    val name2: String,
    val url: String,
    val symbol: String,
    val price: Double, 
    val change: Double, 
    val ratio: Double
)
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
  \begin{lstlisting}[
    language=kotlin,
    basicstyle=\scriptsize\ttfamily,
    numbers=none,
    caption={\texttt{Presentation} class},
    label={lst:presentation-class},
]
data class Presentation(
    val id: Long,
    val title: String, 
    val speakerName: String,
    val summary: String
)
\end{lstlisting}
\end{minipage}
\end{center}

The application's repository contains a list of 10 instances of the
\texttt{Presentation} class and 20 instances of the \texttt{Stock} class. Each
list is used to generate a respective HTML view. Although the instances are
kept in memory, the repository uses the \texttt{Observable} class from the
\textit{RxJava} library to interleave list items with a delay of 1 millisecond.
This delay promotes context switching and frees up the calling thread to handle
other requests in non-blocking scenarios, mimicking actual I/O operations.

By using the \texttt{blockingIterable} method of the \texttt{Observable} class,
we provide a blocking interface for template engines that do not support
asynchronous data models, while still simulating the asynchronous nature of the
data source to enable PSSR\@. Template engines that do not support non-blocking
I/O for PSSR include KotlinX, Rocker, JStachio, Pebble, Freemarker, Trimou, and
Velocity. HtmlFlow supports non-blocking I/O through suspendable templates and
asynchronous rendering, while Thymeleaf enables it using the
\texttt{ReactiveDataDriverContextVariable} in conjunction with a non-blocking
Spring \texttt{ViewResolver}.

The aforementioned blocking template engines are used in the context of Virtual
Threads or alternative coroutine dispatchers, allowing the handler thread to be
released and reused for other requests.

The Spring WebFlux core implementation uses Project Reactor to support a
reactive programming model: each method returns a \texttt{Flux<String>} as the
response body, which acts as a publisher that progressively streams the HTML
content to the client. The implementation includes four main approaches to
PSSR\@:
\begin{itemize}
  \item \textbf{Reactive}: The template engine is used in a reactive context,
        where the HTML content rendered using the reactive programming model.
        An example of this approach is the Thymeleaf template engine
        when using the \texttt{ReactiveDataDriverContextVariable} in conjunction with a
        non-blocking Spring \texttt{ViewResolver}.
  \item \textbf{Virtual}: The template engine is used in a non-blocking
        context, where the HTML content is rendered within the context of Virtual
        Threads. This method is used for the template engines that do not traditionally
        support non-blocking I/O, be it either because they use external DSLs and in
        consequence only support the blocking \texttt{Iterable} interface, or
        because they do not support the asynchronous rendering of HTML content.
  \item \textbf{Suspendable}: The template engine is used in a suspendable
        context, where the HTML content is rendered within the context of a suspending function.
        An example of this approach is the HtmlFlow template engine, which
        supports suspendable templates with the use of the \texttt{Flow} class from the
        Kotlin standard library.
\end{itemize}
% It also includes methods using Kotlin coroutines and other
% asynchronous mechanisms supported by template engines, such as the
% \texttt{writeAsync} method from HtmlFlow, which enables non-blocking I/O
% using \textit{continuations}. 
The Spring MVC implementation uses handlers based solely on the blocking
interface of the \texttt{Observable} class. To enable PSSR in this context, we
utilize the \texttt{StreamingResponseBody} interface, which allows the
application to write directly to the response \texttt{OutputStream} without
blocking the servlet container thread. According to the Spring documentation,
this class is \textit{a controller method return value type for asynchronous
request processing where the application can write directly to the response
OutputStream without holding up the Servlet container thread.}

In Spring MVC, \texttt{StreamingResponseBody} enables asynchronous writing
relative to the request-handling thread, but the underlying I/O remains
blocking—specifically the writes to the \texttt{OutputStream}. When using
Virtual Threads, the I/O operations are more efficient when compared to
platform threads, as they are executed in the context of a lightweight thread.
Most of the computation is done in a separate thread from the one that receives
each request; we use a thread pool \texttt{TaskExecutor} to process requests,
allowing the application to scale and handle multiple clients more efficiently
as opposed to the default \texttt{TaskExecutor} implementation, which tries to
create a thread for each request.

However, the Spring MVC implementation does not effectively support PSSR for
these templates, as HTML content is not streamed progressively to the client.
This is because the response is only sent once the content written to the
\texttt{OutputStream} exceeds the output buffer size, which defaults to 8KB\@.
As a result, the client receives the response only after the entire HTML
content is rendered, defeating the purpose of PSSR in this context.

This implementation includes two main approaches to PSSR\@:
\begin{itemize}
      \item \textbf{Blocking}: The template engine is used in a blocking context,
            where the HTML content is rendered using the blocking interface of the
            \texttt{Observable} class.
      \item \textbf{Virtual}: The template engine is used in a non-blocking context,
            where the HTML content is rendered within the context of Virtual Threads.
\end{itemize}

The Quarkus implementation also uses handlers based on the blocking interface
of the \texttt{Observable} class. It implements the \texttt{StreamingOutput}
interface from the JAX-RS specification to enable PSSR, allowing HTML content
to be streamed to the client. While \texttt{StreamingOutput} also uses blocking
I/O, it operates on Vert.x worker threads, which prevents blocking of the event
loop. When Virtual Threads are used, the I/O operations are handled
efficiently, as they are executed in lightweight threads.

The Quarkus implementation supports PSSR for these templates by configuring the
response buffer size in the \textit{application.properties} file. The default
buffer size is 8KB, but we reduced it to 512 bytes, which allows the response
to be sent to the client progressively as the HTML content is rendered.

This implementation includes three main approaches to PSSR\@:
\begin{itemize}
      \item \textbf{Blocking}: The template engine is used in a blocking context,
            where the HTML content is rendered using the blocking interface of the
            \texttt{Observable} class.
      \item \textbf{Virtual}: The template engine is used in a non-blocking context,
            where the HTML content is rendered within the context of Virtual Threads.
      \item \textbf{Reactive}: The template engine is used in a reactive context,
            where the HTML content is rendered using the reactive programming model.
            In this case, we use the HtmlFlow template engine, which
            supports asynchronous rendering through the \texttt{writeAsync} method.
\end{itemize}