### 2025-05-29 docs/article/Sections/conclusion.tex
Completed Article disclaimers

\section{Conclusion}

In recent decades, non-blocking I/O has become the standard approach for building
highly responsive and scalable web servers. However, traditional synchronous
programming models are not compatible with non-blocking APIs, which typically
rely on callback-based conventions such as
\textit{continuation-passing style} (CPS)~\cite{scheme} or
\textit{promises}~\cite{promise}. These approaches not only hinder
sequential readability but also increase code verbosity, making them more
error-prone.
Alternatives like the \texttt{async}/\texttt{await} idiom~\cite{async_await}
or \textit{suspend functions}~\cite{elizarov2021coroutines} simplify
asynchronous programming by mimicking a sequential style without blocking
threads.
Recent proposals~\cite{carvalho2023async,wise2024pssr} have applied contemporary
asynchronous idioms to SSR web templates, demonstrating how they can overcome
the scalability bottlenecks present in traditional web template engines.

As an alternative, Java Virtual Threads can be applied to any blocking I/O call,
leveraging the Java runtime to transparently intercept and convert it into non-blocking I/O,
without requiring any changes to the calling code or exposing its internal complexities.
In this work, we explored how this technique can be applied to traditional
web template engines and whether it can achieve performance competitive with
reactive approaches provided by frameworks such as Thymeleaf~\cite{webflux} and HtmlFlow~\cite{htmlflow}.
Our benchmarks across Spring WebFlux, Spring MVC, and Quarkus show that synchronous
non-blocking execution using virtual threads consistently delivers performance
comparable to asynchronous non-blocking approaches under high concurrency.
These findings highlight virtual threads as a promising alternative to complex
asynchronous programming models, offering a simpler development experience
without compromising scalability or responsiveness.