### 2025-07-13 docs/article/Sections/introduction.tex

\section{Introduction}

% Falta adicionar referências

Modern web applications rely on different rendering strategies to optimize
performance, user experience, and \textit{scalability}. The two most dominant
approaches are \textit{Server-Side Rendering (SSR)} and \textit{Client-Side
    Rendering (CSR)}. SSR generates HTML content on the server before sending it to
the client, resulting in a faster \textit{First Contentful Paint (FCP)}
~\cite{Edgar2024-FCP} and better Search Engine Optimization (SEO). However, SSR
can increase server load and reduce \textit{throughput}\footnote{The number of
    requests the server can handle per second (RPS)} since each request requires
additional processing before responding. In contrast, CSR shifts the rendering
workload to the browser—the server initially sends a minimal HTML document with
JavaScript, which dynamically loads the page content. While CSR reduces the
server’s burden, it can lead to a slower \textit{FCP}, as users must wait for
JavaScript execution before meaningful content appears.

\textit{Progressive Server-Side Rendering (PSSR)} combines benefits from both SSR and CSR
by streaming HTML content progressively. This technique enhances user-perceived performance by
allowing progressive rendering as data becomes available, significantly reducing the
\textit{Time to First Byte (TTFB)} and improving perceived load times compared to traditional
SSR approaches~\cite{wise2024pssr}. In this respect,
PSSR is similar to CSR in that the server initially sends a minimal HTML
document to the client and subsequently streams additional HTML fragments.
However, unlike CSR, PSSR retains all rendering responsibilities on the server side,
thereby reducing the load on the client. As a result, the client does
not need to execute JavaScript or make additional requests to retrieve page content.
The streaming nature of PSSR allows users to see content progressively as it becomes
available, rather than waiting for the complete page to be rendered server-side,
thus providing a more responsive user experience with measurably lower TTFB values.

\textit{Low-thread} servers, also known as
\textit{event-driven}~\cite{event-driven-servers}, have gained prominence in
contemporary web applications due to their ability to efficiently manage a large
number of concurrent I/O operations with minimal resources, thus promoting
better scalability.
By leveraging asynchronous I/O operations—such as
database queries and API calls, servers can avoid blocking threads while
waiting for data, thereby maximizing throughput and responsiveness. To support
this non-blocking architecture, PSSR implementations require template engines
that are compatible with asynchronous data models~\cite{carvalho2023async}.
Some modern template
engines, such as HtmlFlow~\cite{htmlflow} and Thymeleaf~\cite{thymeleaf}, have
been designed with these capabilities in mind. However, many legacy
template engines—particularly those using external domain-specific languages
(DSLs)~\cite{Fowler03}—still depend on blocking interfaces like
\texttt{Iterable} for data processing. This blocking behavior forces server
threads to remain idle until the entire HTML output is ready, undermining the
performance benefits of non-blocking I/O and limiting scalability in
high-concurrency environments.

With the introduction of \textit{Virtual
    Threads}~\footnote{\url{https://openjdk.org/jeps/444}} in Java 21, it is now
possible to execute blocking I/O operations in a scalable, lightweight manner.
This capability allows legacy template engines—often reliant on blocking
interfaces—to operate efficiently in high-concurrency, non-blocking
environments without requiring complex asynchronous programming models. As a
result, PSSR can now be implemented using familiar HTML templates, simplifying
development and improving maintainability.

This work explores the current landscape of non-blocking PSSR, focusing on two
primary paradigms: \textit{reactive programming} and \textit{coroutines}, both
of which have been used to achieve asynchronous I/O in the Java ecosystem. As
an alternative, we investigate whether Java’s Virtual Threads can offer
comparable performance while preserving the simplicity of synchronous code.
Section 2 reviews the state-of-the-art in PSSR and template engine design.
Section 3 outlines the limitations of conventional engines in asynchronous
settings and presents our proposed approach. Section 4 details the benchmark
methodology, followed by the results and analysis in Section 5. Conclusions are
presented in Section 6.