### 2025-07-13 docs/article/Sections/state-of-the-art-web-servers.tex

\subsection{Web Framework Architectures and Approaches to PSSR}
\label{sec:web-frameworks}

In traditional thread-per-request architectures, each incoming request is
handled by a dedicated thread. The web server maintains a thread pool from
which a thread is assigned to handle each request. As the load increases, the
number of active threads can grow rapidly, potentially exhausting the pool.
This can lead to performance and scalability issues, as the system may become
bogged down by context switching and thread management
overhead~\cite{kant2000scalable}. In the Java ecosystem, Spring MVC is one of
the most widely used frameworks that follow this model, according to several
reports such as the Stack Overflow 2024 Developer
Survey~\footnote{~\url{https://survey.stackoverflow.co/2024/technology\#1-web-frameworks-and-technologies}}.

On the other hand, in modern \textit{low-thread} architectures, the server uses
a small number of threads to handle a large number of requests.
\textit{Low-thread} servers, also known as
\textit{event-driven}~\cite{event-driven-servers}, offer a significant
advantage in efficiently managing a high number of concurrent I/O operations
with minimal resource usage.

The \textit{non-blocking} I/O model employed in low-thread servers is
well-suited for handling large volumes of data asynchronously~\cite{Meijer12}.
This combination of low-thread servers and asynchronous data models has
facilitated the development of highly scalable, responsive, and resilient web
applications capable of managing substantial data loads~\cite{Jin15}. The
prominence of this concept increased with the advent of Node.js in 2009, and
subsequently, various technologies adopted this approach in the Java ecosystem,
including Netty, Akka HTTP, Vert.X, and Spring WebFlux.
% Among these, Spring WebFlux stands out as the most widely used web framework in
% Java, according to surveys like JetBrains' State of Java report (2021) and
% community metrics from platforms like Github and StackOverflow.
The \textit{non-blocking} I/O model in low-thread servers functions optimally
only when HTTP handlers avoid blocking. Therefore, HTML templates need to be
proficient in dealing with the asynchronous APIs provided by data models. While
most legacy web templates struggle with asynchronous models, DSLs for HTML face
no such limitations, leveraging all constructions available in the host
programming language. However, the unexpected intertwining of asynchronous
handlers' completion and HTML builders' execution may potentially lead to
malformed HTML with an unexpected layout~\cite{wise2024pssr}.

An alternative approach involves utilizing user-level threads, while
maintaining a blocking I/O and a synchronous programming paradigm. However,
this approach still requires a user-level I/O subsystem capable of mitigating
system-level blocking, which is crucial for the performance of I/O-intensive
applications. This technique offers a lightweight solution for efficiently
managing a larger number of concurrent sessions by minimizing per-thread
overhead. In 2020, Karsten~\cite{karsten2020} demonstrated how this strategy
supports a synchronous programming style, effectively abstracting away the
complexities associated with managing continuations in asynchronous
programming.

Among mainstream technologies, the Kotlin programming language introduces a new
abstraction for managing coroutines and provides structured concurrency, which
ensures that coroutines are \textit{scoped}, \textit{cancellable}, and
\textit{coordinated} within a well-defined
lifecycle~\cite{elizarov2021coroutines}. Although this model embraces the
\texttt{async}/\texttt{await} feature~\cite{async_await}, enabling non-blocking
routines to mimic the structure of synchronous ones, it still lacks a
user-level I/O subsystem and relies on an explicit I/O dispatcher tied to a
dedicated thread pool that frees worker threads for handling processing tasks.
Moreover, web templates using their own templating dialects, such as JSP,
Thymeleaf, or Handlebars, are unable to take advantage of such constructs, as
they typically support only a limited subset of the host library's API—most
commonly just the \texttt{Iterable} interface.

Only Java virtual threads, introduced in
JEP~444\footnote{\url{https://openjdk.org/jeps/444}}, adopt a similar approach
to Karsten's proposal, using user-mode threads to preserve a synchronous
programming model while still allowing blocking I/O operations without tying up
platform threads. When a virtual thread performs a blocking I/O operation, the
JVM intercepts the call and transparently parks the virtual thread, freeing the
underlying platform thread to perform other tasks. Once the I/O operation
completes, the virtual thread is rescheduled on a platform thread and resumes
execution. This mechanism is enabled by the JVM's integration with non-blocking
system calls at the OS level, allowing developers to write traditional
synchronous code while benefiting from the scalability of non-blocking I/O.

The primary advantage of virtual threads lies in their ability to eliminate the
need for complex asynchronous programming models. Unlike reactive or structured
concurrency approaches that require developers to adopt complex paradigms, virtual
threads enable non-blocking execution of existing synchronous APIs without code
modification. This significantly reduces the learning curve and maintenance
burden associated with asynchronous programming while maintaining comparable
performance characteristics.

\subsubsection{Spring MVC}

Spring MVC is a synchronous web framework built on the thread-per-request
model. When an HTTP request is received, it is handled by a dedicated thread
from the server’s thread pool (e.g., Tomcat, Jetty, or Undertow). This thread
is responsible for executing the full request lifecycle synchronously,
including invoking controllers, executing business logic, and rendering the
response view. Because the model is blocking, the thread remains occupied
throughout the request—even during I/O operations such as database queries or
external API calls—which can lead to performance bottlenecks under high
concurrency.

Controllers in Spring MVC typically return Java objects or models that are
processed by view resolvers in conjunction with template engines like JSP,
Thymeleaf, or FreeMarker. These templates generate the full HTML response in
one pass, which is then sent to the client only after all necessary data has
been gathered and rendered. By default, this architecture limits the ability to
perform PSSR, as traditional rendering waits for all data before sending any
output. This results in higher latency for clients and less responsive page
loads, particularly for data-intensive views. However, Spring MVC provides a
mechanism for response streaming via the
\texttt{StreamingResponseBody}\footnote{\url{https://docs.spring.io/spring-framework/docs/current/javadoc-api/org/springframework/web/servlet/mvc/method/annotation/StreamingResponseBody.html}}
interface, introduced in Spring 4.2. When a controller method returns a
\texttt{StreamingResponseBody}, Spring writes directly to the
\texttt{HttpServletResponse} output stream, allowing the server to send parts
of the response incrementally as they become available. This is particularly
useful for:
\begin{itemize}
    \item Sending large responses without buffering the entire output in memory.
    \item Writing dynamic HTML in fragments as data is fetched or computed.
    \item Reducing Time to First Byte (TTFB) and improving perceived performance.
\end{itemize}

\begin{lstlisting}[
    language=Java,
    basicstyle=\scriptsize\ttfamily,
    numbers=none,
    caption={StreamingResponseBody handler in Spring MVC},
    label={lst:streaming-response-body}
]
  @GetMapping("/stream")
  fun handleStream(): StreamingResponseBody? {
      return StreamingResponseBody { outputStream ->
          for (i in 0..9) {
              val htmlFragment = "<p>Chunk " + i + "</p>\n"
              outputStream.write(htmlFragment.toByteArray())
              outputStream.flush() // ensure partial response is sent
              Thread.sleep(500) // simulate delay
          }
      }
  }
\end{lstlisting}

In the example shown in \autoref{lst:streaming-response-body}, HTML content is
written and flushed in discrete chunks, allowing the client to progressively
render the response as it arrives, while the server continues processing
additional data. However, this technique has two limitations. First, it does
not eliminate the blocking nature inherent in Spring MVC, as the handling
thread remains active during the streaming process. Second, it is constrained
by the servlet response buffer. By default, most servlet containers (e.g.,
Tomcat) use a response buffer size of approximately 8KB. Data written to the
response output stream is held in this buffer until it reaches capacity.
Consequently, if the total size of the initial HTML content is less than the
buffer threshold, the server will not transmit any data to the client until the
buffer is filled. This buffering behavior introduces a delay in sending the
first chunk of content, thereby negating the benefits of progressive rendering.
In scenarios where the rendered HTML template does not produce sufficient data
to exceed the buffer threshold early, PSSR becomes ineffective, and the client
experiences a delayed initial response.

\subsubsection{Spring WebFlux}

Spring WebFlux is a reactive web framework built on a non-blocking,
event-driven architecture that enables efficient handling of a large number of
concurrent connections with minimal resource usage. Unlike traditional
servlet-based stacks that rely on a thread-per-request model, WebFlux operates
on reactive runtimes like Netty, where a small, fixed-size event loop manages
all I/O events. Incoming HTTP requests are routed to handler functions or
annotated controllers, which return reactive types such as \texttt{Mono<T>} and
\texttt{Flux<T>}—\texttt{Publisher} implementations provided by the Project
Reactor library~\cite{projectreactor}. These implementations adhere to the
Reactive Streams specification~\cite{ReactiveStreams}, enabling
interoperability with other compliant reactive stream libraries. A
\texttt{Mono} represents a single asynchronous value (or none)~\cite{promise},
while a \texttt{Flux} represents a stream of zero or more
values~\cite{rx-observable}. These abstractions allow the entire request
lifecycle—from controller invocation to response rendering—to be orchestrated
as a chain of non-blocking, asynchronous operations. I/O-bound tasks such as
database queries or external API calls do not occupy threads during execution,
enabling the framework to scale gracefully under high load and maintain
responsiveness even in resource-constrained environments.

PSSR is natively supported in WebFlux via \texttt{Flux}-based controllers. When
a controller returns a \texttt{Flux<String>} or another stream of HTML
fragments, WebFlux can begin streaming content to the client as soon as the
first elements are emitted.
% This allows browsers to incrementally render parts of the response without waiting
% for the entire payload, significantly reducing Time to First Byte (TTFB) and
% improving user-perceived performance. By integrating PSSR directly into the
% reactive processing pipeline, WebFlux facilitates responsive and resilient web
% applications well-suited for modern microservices architectures.

\begin{lstlisting}[
    language=Java,
    basicstyle=\scriptsize\ttfamily,
    numbers=none,
    caption={Progressive Server-Side Rendering in Spring WebFlux},
    label={lst:pssr-webflux}
]
@GetMapping(value = "/pssr", produces = MediaType.TEXT_HTML_VALUE)
public Flux<String> renderChunks() {
    return Flux.range(1, 10)
               .delayElements(Duration.ofMillis(500))
               .map(i -> "<p>Chunk " + i + "</p>\n");
}
\end{lstlisting}

\autoref{lst:pssr-webflux} shows a method that emits one HTML chunk every 500 milliseconds. As
each chunk is generated, it is immediately written to the HTTP response,
allowing the client to begin rendering content even as more data is still being
produced. Unlike \texttt{StreamingResponseBody} in Spring MVC, which is
constrained by servlet buffer thresholds and blocking semantics, WebFlux
ensures that each emitted item is pushed to the client as soon as the network
is ready, without occupying a dedicated thread. Moreover, WebFlux integrates
seamlessly with reactive data sources such as R2DBC (Reactive Relational
Database Connectivity) and reactive NoSQL drivers, making it possible to build
end-to-end non-blocking applications. HTML builders and DSLs used in
conjunction with WebFlux must be designed to accommodate reactive streams,
ensuring that view generation does not violate the non-blocking contract by
performing blocking I/O or awaiting asynchronous operations imperatively.

\subsubsection{Quarkus}

Quarkus is a modern, cloud-native Java framework that supports both imperative
and reactive programming models. It is built on top of Vert.X, which provides a
non-blocking, event-driven runtime similar to Netty. Incoming HTTP requests are
processed asynchronously on a small, fixed-size event loop thread pool,
allowing the framework to handle many concurrent connections efficiently
without blocking threads. Quarkus provides support for reactive APIs like
Mutiny~\cite{mutiny2021}, a modern alternative to Spring
Reactor~\cite{projectreactor}. Like Reactor, Mutiny complies with the Reactive
Streams specification~\cite{ReactiveStreams}. For traditional imperative-style
request handling, Quarkus provides integration with
JAX-RS~\cite{burke2013restful}, where resource methods can return synchronous
types or asynchronous constructs such as \texttt{CompletionStage}. To enable
PSSR in imperative endpoints, Quarkus leverages the JAX-RS
\texttt{StreamingOutput} interface. By returning a \texttt{StreamingOutput}
instance, developers can write parts of the HTTP response body incrementally as
data becomes available. This approach allows the server to flush HTML fragments
progressively to the client, improving the perceived responsiveness especially
for long-running or large responses. Like other streaming mechanisms
based on servlet buffers, the initial data flush may be delayed until internal
buffers (commonly around 8KB) are filled. However, Quarkus allows configuring
the buffer size to a smaller value, which can help mitigate this issue.

\begin{lstlisting}[
  language=java,
  basicstyle=\scriptsize\ttfamily,
  numbers=none,
  caption={Progressive Server-Side Rendering in Quarkus using StreamingOutput},
  label={lst:pssr-quarkus}
]
@GET
@Path("/stream")
@Produces(MediaType.TEXT_HTML)
public StreamingOutput streamHtml() {
    return output -> {
        PrintWriter writer = new PrintWriter(output);
        for (int i = 1; i <= 5; i++) {
            writer.println("<p>Chunk " + i + "</p>");
            writer.flush(); // Flush each chunk incrementally
            Thread.sleep(500); // Simulate processing delay
        }
        writer.flush();
    };
}
\end{lstlisting}

\autoref{lst:pssr-quarkus} shows how \texttt{StreamingOutput} can be used to send
partial HTML content in chunks, allowing browsers to progressively render the
response while the server continues processing. Quarkus’ architecture ensures
that the underlying event loop threads are not blocked by these writes, as the
framework offloads blocking I/O operations to worker threads. This combination
of non-blocking event-driven runtime and incremental response streaming enables
efficient PSSR even within imperative programming styles.