### 2025-07-29 docs/report/rfG48.tex

% Classe do documento e parâmetros gerais.
\documentclass[a4paper,openright,twoside,11pt]{report}

% Packages utilizadas e respetivos parâmetros.
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\addto{\captionsenglish}{\renewcommand{\bibname}{References}}
\addto{\captionsenglish}{\renewcommand{\contentsname}{Contents}}
\usepackage{lipsum} % gerador de texto
\usepackage{graphicx}
\usepackage{url}
\usepackage[Algoritmo]{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[dvipsnames]{xcolor}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{listingstyles}
\usepackage{amssymb}
\usepackage{bookmark}

\renewcommand{\algorithmicrequire}{\textbf{Dados: }}
\renewcommand{\algorithmicensure}{\textbf{Resultado: }}

% Definições das dimensões das páginas
\setlength{\textheight}{24.00cm}
\setlength{\textwidth}{15.50cm}
\setlength{\topmargin}{0.35cm}
\setlength{\headheight}{0cm}
\setlength{\headsep}{0cm}
\setlength{\oddsidemargin}{0.25cm}
\setlength{\evensidemargin}{0.25cm}

%\renewcommand{\baselinestretch}{1}

% Página inicial (capa)
\title{
   \vspace{-50mm}
   \begin{minipage}[l]{\textwidth}
      \hspace{-20mm}\resizebox{75mm}{!}{\includegraphics{./figures/logoISELnew2.png}}\\
   \end{minipage}\\[20mm]
   {\fontsize{22pt}{26pt}\selectfont\textbf{Non-Blocking Progressive Server-Side Rendering Benchmark}}\\[10mm]
   {\fontsize{14pt}{18pt}\selectfont\textbf{A Performance Comparison of Progressive SSR Techniques}}\\[5mm]
}

% Nome dos autores (um por linha)
\author{
\begin{tabular}{ll}
             & Bernardo Pereira \\[50mm]
\end{tabular}}

\date{
\begin{tabular}{ll}
  {Supervisor:} & Miguel Carvalho, ISEL \\
\end{tabular}\\[10mm]
% Deixar o indicador respetivo em função da versão do relatório.
Final Report written for Project and Seminar\\
BSc in Computer Science and Engineering\\[20mm]
June 2025}

\begin{document}
\pagenumbering{roman}
\thispagestyle{empty}
\maketitle

\baselineskip 18pt % line spacing: 12pt for single, 18pt for 1 1/2, and 24pt for double spacing

\newpage
\thispagestyle{empty}
% Fim da contracapa

% Página com identificação completa (número e nome) e assinaturas do(s) estudante(s) e do(s) orientador(es)
\cleardoublepage
\setcounter{page}{1}
\begin{center}
  \textsc{\LARGE Instituto Superior de Engenharia de Lisboa}\\[30mm]

  {\fontsize{22pt}{26pt}\selectfont\textbf{Non-Blocking Progressive Server-Side Rendering Benchmark}}\\[10mm]
  {\fontsize{14pt}{18pt}\selectfont\textbf{A Performance Comparison of Progressive SSR Techniques}}\\[30mm]

  \begin{tabular}{rl}
    50493 & Bernardo André Batista Correia Murta Pereira \\[10mm]
          & \rule{75mm}{0.5pt}                           \\[5mm]
  \end{tabular}\\[10mm]

  \begin{tabular}{rl}
    Supervisor: & Miguel Carvalho, ISEL \\[10mm]
                & \rule{75mm}{0.5pt}    \\[5mm]
  \end{tabular}\\[10mm]

  Final report written for Project and Seminar\\ BSc in Computer Science and Engineering\\[20mm]
  June 2025\\
\end{center}

% Página de resumo em Português
\cleardoublepage
\chapter*{Abstract}

Modern web applications increasingly demand rendering techniques that optimize performance, responsiveness, and scalability. Progressive Server-Side Rendering (PSSR) bridges the gap between Server-Side Rendering and Client-Side Rendering by progressively streaming HTML content, improving perceived load times.

However, traditional HTML template engines often rely on blocking interfaces
that hinder their use in asynchronous, non-blocking contexts required for PSSR.

This work explores how Java Virtual Threads, introduced in Java 21, enable
non-blocking execution of blocking I/O operations, allowing the reuse of
traditional template engines for PSSR without complex asynchronous programming
models. We benchmark multiple engines across Spring WebFlux, Spring MVC, and
Quarkus using reactive, suspendable, and virtual-thread-based approaches.
Results show that virtual threads allow blocking engines to scale comparably to
those designed for non-blocking I/O, achieving high throughput and
responsiveness under load. This demonstrates that Virtual Threads provide a
compelling path to simplify the implementation of PSSR with familiar HTML
templates, significantly lowering the barrier to entry while maintaining
performance.

\vspace{5mm}

{\bf Keywords:} Web Templates; Server-Side Rendering; Non-Blocking; Java Virtual Threads; Asynchronous.

%% Página de agradecimentos
%\cleardoublepage
%\chapter*{Agradecimentos}
%Texto dos agradecimentos. É opcional.\\

% Geração do índice de conteúdos
\cleardoublepage
\tableofcontents \cleardoublepage

% Geração do índice de figuras e de tabelas
%\listoffigures \cleardoublepage
%\listoftables \cleardoublepage

% Reiniciar a numeração de páginas
\setcounter{page}{1}
\pagenumbering{arabic}

% Capitulo 1
\include{Sections/introduction}

% Capitulo 2
\include{Sections/state-of-the-art}

% Capitulo 3
\include{Sections/problem-statement}

% Capitulo 4
\include{Sections/benchmark}

% Capitulo 5
\include{Sections/results}

% Capitulo 6
\include{Sections/conclusion}

% Referências
\bibliographystyle{unsrt}
\bibliography{bibliography}
\addcontentsline{toc}{chapter}{References}

\end{document}